---
title: "Metric Analysis"
format:
  html:
    fig-dpi: 200
    code-fold: true
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: false
---

```{r}
#| label: prep
#| cache: false
#| output: false
options(scipen = 999)

pacman::p_load(
  conflicted,
  dplyr,
  htmltools,
  stringr,
  tidyr,
  reactable,
  purrr,
  broom,
  knitr,
  kableExtra,
  tictoc,
  furrr,
  parallelly,
  splm,
  spdep,
  spatialreg,
  sf,
  skimr,
  ggplot2,
  ggpubr,
  readr,
  lmtest,
  sandwich,
  performance,
  snakecase,
  ggtext
)

pacman::p_load_gh(
  'Food-Systems-Research-Institute/SMdata',
  'Food-Systems-Research-Institute/SMdocs'
)

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::pull(),
  stats::lag(),
  base::setdiff(),
  .quiet = TRUE
)

dp_meta <- readRDS(here::here("0_prep", "output", "dp_meta.rds"))
dp_weights <- readRDS(here::here("0_prep", "output", "dp_weights.rds"))
dp_metrics_county_adj <- readRDS(here::here("0_prep", "output", "dp_metrics_county_adj.rds"))
dp_metrics_state_adj <- readRDS(here::here("0_prep", "output", "dp_metrics_state_adj.rds"))
```

```{r}
#| label: graph_functions
#| output: false
#| echo: false

# Take formatted figure data, a palette for text colors, the label dimension
# lookup from prepare_figure_data(). Return ggplot graph object
pct_change_graph <- function(figure_data, 
                             dp_text_palette, 
                             label_dimension_lookup) {

  # Renaming input to preserve it for error checking
  data <- figure_data
  
  # Standardize column names
  if ("conf_low" %in% names(data) && "conf_high" %in% names(data)) {
    data <- data %>%
      rename(ci_low = conf_low, ci_high = conf_high)
  }

  data <- data %>%
    mutate(across(c(estimate, ci_low, ci_high), ~ as.numeric(.x))) %>%
    mutate(
      across(c(estimate, starts_with('ci|conf')), as.numeric),
      outcome = ifelse(is.na(outcome), 'no change', outcome),
      outcome = snakecase::to_title_case(outcome)
    )

  ggplot(data, aes(x = estimate, y = combined_label)) +
    geom_point(aes(color = outcome)) +
    geom_errorbarh(aes(xmin = ci_low, xmax = ci_high, color = outcome)) +
    geom_vline(xintercept = 0) +
    
    # Invisible points to create a second legend (dimension colors)
    geom_point(
      aes(fill = dimension, color = dimension), 
      shape = 22, 
      alpha = 0, 
      show.legend = TRUE
    ) +
    
    # Add stars
    geom_text(
      aes(x = ci_high + 0.1, label = sig),
      hjust = 0,
      size = 4
    ) +
    
    # Outcome legend (color of real points and lines)
    scale_color_manual(
      values = c('Better' = 'darkgreen', 'No Change' = 'grey', 'Worse' = 'purple'),
      name = "Outcome"
    ) +
    
    # Dimension legend (fill colors)
    scale_fill_manual(
      values = dp_text_palette,
      name = "Dimension",
      labels = snakecase::to_title_case(names(dp_text_palette))
    ) +
    
    # Add both legends
    guides(
      color = guide_legend(order = 1, nrow = 2),
      fill = guide_legend(
        override.aes = list(shape = 22, size = 4, alpha = 1, color = dp_text_palette),
        order = 2,
        nrow = 2
      )
    ) +
    
    theme_minimal() +
    
    labs(
      x = 'Average Percent Change per Year',
      y = 'Indicator with Metric'
    ) + 
    
    scale_y_discrete(
      labels = make_y_axis_labels(label_dimension_lookup, dp_text_palette)
    ) +
    
    theme(
      legend.position = "top",
      legend.direction = "horizontal",
      legend.title.position = "top",
      text = element_text(family = "Times New Roman"),
      axis.title = element_text(family = "Times New Roman"),
      axis.text = element_text(family = "Times New Roman"),
      legend.title = element_text(family = "Times New Roman", hjust = 0.5),
      legend.text = element_text(family = "Times New Roman"),
      axis.text.y.left = ggtext::element_markdown(),
    )
}
```

This page explores time series data at the state and county level, then runs some regressions to see whether metrics are improving or declining over time at the county level. We first run weighted OLS models with robust SEs. Then we use spatial error models weighted by both geography and weighting variables.

## Time Series

Plot time series of each metric that has at least two data points. These are unweighted values. LOESS curves are shown in blue, while individual county/state data are shown in black

### County Data

First, we prepare a vector names of our time series variables for counties:

```{r}
#| label: prep_series
#| output: false
# Prepare vector of county variables that are time series
str(dp_metrics_county_adj)
vars <- unique(dp_metrics_county_adj$variable_name)
series_vars <- map(vars, ~ {
  count <- dp_metrics_county_adj %>% 
    filter(variable_name == .x) %>% 
    pull(year) %>% 
    unique() %>% 
    length()
  if (count > 1) {
    return(.x)
  } else {
    return(NULL)
  }
}) %>% 
  keep(~ !is.null(.x)) %>%
  unlist()

dir.create('4_trends/output/trend_plots', recursive = TRUE, showWarnings = FALSE)
saveRDS(series_vars, '4_trends/output/series_vars.rds')
```

Now we plot each graph individually and arrange them together:

```{r}
#| label: county_time_series
#| fig-height: 25
#| fig-width: 12
#| cache: true
plots <- map(series_vars, ~ {
  df <- dp_metrics_county_adj %>%
    filter(variable_name == .x) %>%
    ggplot(aes(x = year, y = value)) +
    geom_line(
      aes(group = fips),
      alpha = 0.3
    ) +
    geom_smooth(
      method = 'loess',
      lwd = 2
    ) +
    labs(
      x = 'Year',
      title = .x
    ) +
    theme_classic()
})

ggarrange(
  plotlist = plots,
  ncol = 4,
  nrow = 11
)
```

### State Data

Again, start by getting a vector of time series metrics for state data, then plot all metrics:

```{r}
#| label: state_time_series
#| fig-height: 12
#| fig-width: 12
vars <- unique(dp_metrics_state_adj$variable_name)
state_series_vars <- map(vars, ~ {
  count <- dp_metrics_state_adj %>% 
    filter(variable_name == .x) %>% 
    pull(year) %>% 
    unique() %>% 
    length()
  if (count > 1) {
    return(.x)
  } else {
    return(NULL)
  }
}) %>% 
  keep(~ !is.null(.x)) %>%
  unlist()
saveRDS(state_series_vars, '4_trends/output/state_series_vars.rds')

plots <- map(state_series_vars, ~ {
  df <- dp_metrics_state_adj %>%
    filter(variable_name == .x) %>%
    mutate(across(c(year, value), as.numeric)) %>% 
    ggplot(aes(x = year, y = value)) +
    geom_line(
      aes(group = fips),
      alpha = 0.3
    ) +
    geom_smooth(
      method = 'loess',
      lwd = 2
    ) +
    labs(
      x = 'Year',
      title = .x
    ) +
    theme_classic()
})

ggarrange(
  plotlist = plots,
  ncol = 3,
  nrow = 5
)
```

### Tiny Trend Graphs

Outputs not shown here - these will be added to a table in the paper.

```{r}
all_series_vars <- c(series_vars, state_series_vars)

# Function to get tiny plots from county and state vars
get_tiny_plots <- function(df, vars) {
  map(vars, ~ {
    df <- df %>%
      filter(variable_name == .x, !is.na(value)) %>% 
      ggplot(aes(x = year, y = value)) +
      geom_line(
        aes(group = fips),
        alpha = 0.25,
        color = 'grey',
        linewidth = 0.5
      ) +
      geom_smooth(
        method = "loess", 
        se = FALSE, 
        linewdith = 0.75
      ) +
      coord_cartesian(xlim = c(2000, 2024)) +
      scale_x_discrete(expand = c(0, 0)) +
      scale_y_discrete(expand = c(0, 0)) +
      theme_bw() +
      theme(
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        plot.margin = margin(rep(0.5, 4))
        
      )
  }) %>% 
    setNames(c(vars))
}

# Plots for counties
county_plots <- dp_metrics_county_adj %>% 
  get_tiny_plots(
    vars = series_vars
  )

# Plots for states
state_plots <- dp_metrics_state_adj %>% 
  get_tiny_plots(
    vars = state_series_vars
  )

# One more plot for CPI - food at home
system_plots <- SMdata::metrics %>% 
  filter(variable_name == 'cpiFoodAtHome') %>% 
  get_tiny_plots(vars = 'cpiFoodAtHome')


# Save them all to outputs with variable names
iwalk(c(county_plots, state_plots, system_plots), \(plot, var) {
  ggsave(
    filename = paste0('4_trends/output/trend_plots/fig_trend_', var, '.png'),
    plot = plot,
    width = 1.5,
    height = 1.5,
    units = 'cm',
    dpi = 300,
    bg = NULL
  )
})
```

## Spatial Regressions {#sec-ols}

```{r}
#| label: normalize
#| output: false
str(dp_metrics_county_adj)

vars <- unique(dp_metrics_county_adj$variable_name)
dp_metrics_county_adj_normed <- map(vars, ~ {
  dp_metrics_county_adj %>%
    filter(variable_name == .x) %>% 
    mutate(across(c(year, value), as.numeric)) %>% 
    drop_na(value) %>% 
    group_by(variable_name) %>% 
    mutate(normed_value = min_max(value)) %>% 
    ungroup()
}) %>% 
  list_rbind()
str(dp_metrics_county_adj_normed)  

dir.create('4_trends/output/', recursive = TRUE, showWarnings = FALSE)
saveRDS(dp_metrics_county_adj_normed, '4_trends/output/dp_metrics_county_adj_normed.rds')
```

Create a weights crosswalk that links `variable_name`s to weighting variables.

```{r}
#| label: weights_crosswalk
weights_crosswalk <- data.frame(
  metric = c(
    'GDP',
    'GDP from agriculture',
    'Total number of producers',
    'Total number of operations',
    'Population (annual)',
    'Population (5 year)',
    'Population (5 year, smooth)',
    'Acres of land',
    'Acres of water',
    'Area under agriculture (CDL)',
    'Number employed in agriculture',
    'Acres operated',
    'Acres operated, smooth',
    'Acres of cropland ag land',
    'Acres of woodland ag land',
    'Acres of pasture ag land',
    'Total farm income',
    'Total sales from commodities'
  ),
  variable_name = c(
    'gdp',
    'gdpFromAg',
    'nProducers',
    'nOperations',
    'populationAnnual',
    'population5Year',
    'population5YearSmooth',
    'landAcres',
    'waterAcres',
    'cdlAgAcres',
    'annualAvgEmplvlNAICS11',
    'acresOperated',
    'acresOperatedSmooth',
    'agLandCropAcres',
    'agLandWoodAcres',
    'agLandPastureAcres',
    'totalFarmIncome',
    'totalSalesCommodities'
  )
)

# Link weight vars to the appropriate variables
weights_crosswalk <- dp_meta %>% 
  select(variable_name, weighting) %>% 
  filter(variable_name %in% dp_metrics_county_adj_normed$variable_name) %>% 
  left_join(weights_crosswalk, by = join_by(weighting == metric)) %>% 
  setNames(c('variable_name', 'weight_metric', 'weight_variable')) %>% 
  select(-weight_metric) %>% 
  na.omit()
head(weights_crosswalk)
```

### Models

Spatial error models with the `spatialreg` package. We are now working with:

$$
\Large y = X\beta + u, u = \lambda Wu + \epsilon,
$$

where $y$ are normalized metric values, $X$ is an $n * k$ matrix, $\beta$ is a vector of predictors, $u$ is the spatial autocorrelation error term, $\lambda$ is the spatial error parameter, $W$ is the county weights matrix, and $\epsilon$ is the error term. 

We run this portion of the code in parallel because it has a pretty long run time. Again, results are saved so it doesn't need to be run again unless something changes.

```{r}
#| label: spatial_reg_all
#| output: false

# Slow. Run only when the saved models are missing
# or older than the step-0 input, else load. Delete .rds to force rerun
sar_path <- '4_trends/output/errorsarlm_outs.rds'
in_path  <- here::here('0_prep', 'output', 'dp_metrics_county_adj.rds')
stale <- !file.exists(sar_path) || file.mtime(sar_path) < file.mtime(in_path)

if (stale) {
  dir.create('4_trends/output', recursive = TRUE, showWarnings = FALSE)

  state_link <- SMdata::fips_key %>%
    filter(str_length(fips) == 2, fips != '00') %>%
    select(fips, state_code)

  dat <- dp_metrics_county_adj %>%
    pivot_wider(
      id_cols = fips:year,
      names_from = variable_name,
      values_from = value,
      # TODO: better way to collapse
      values_fn = ~ mean(.x, na.rm = TRUE)
    ) %>%
    filter(str_detect(fips, '^09', negate = TRUE)) %>%
    mutate(link = str_sub(fips, end = 2)) %>%
    left_join(state_link, by = join_by(link == fips)) %>%
    select(-link) %>%
    mutate(across(!c(fips, state_code), as.numeric))

  vars <- names(dat)[!names(dat) %in% c('fips', 'year', 'state_code')]

  # County contiguity from county geometry
  county_sf <- SMdata::neast_counties_2024 %>%
    select(fips) %>%
    filter(str_detect(fips, '^09', negate = TRUE))
  W_county <- nb2mat(poly2nb(county_sf), style = 'B', zero.policy = TRUE)
  dimnames(W_county) <- list(county_sf$fips, county_sf$fips)

  tic()
  # multicore (fork) if possible, otherwise multisession
  if (parallelly::supportsMulticore()) {
    plan(multicore, workers = 6)
  } else {
    plan(multisession, workers = 6)
  }

  sar_outs <- future_map(vars, \(var) {
    cat('\n\nStarting', var)

    tryCatch({
      # Reduce to relevant variable, remove NAs
      df <- dat %>%
        select(fips, year, !!sym(var), state_code) %>%
        drop_na(!!sym(var)) %>%
        mutate(normed = min_max(!!sym(var)))

      # Regular weights. Unweighted vars are absent from the crosswalk
      weight_var <- weights_crosswalk$weight_variable[weights_crosswalk$variable_name == var]
      has_weight <- length(weight_var) == 1 && !is.na(weight_var)
      if (has_weight) {
        weight_df <- dp_weights %>%
          mutate(
            value = as.numeric(value),
            year = as.numeric(year)
          ) %>%
          rename(
            weight_value = value,
            weight_name = variable_name,
            weight_year = year
          ) %>%
          filter(weight_name == weight_var, str_length(fips) == 5)

        # Join weights back to df
        if (all(is.na(weight_df$weight_year))) {
          df <- df %>%
            inner_join(weight_df, by = join_by(fips))
        } else {
          df <- df %>%
            inner_join(weight_df, by = join_by(fips, year == weight_year))
        }

        cat('\nJoined to weights:', weight_var)
      } else {
        cat('\nNo weighting variable')
      }

      model_df <- df

      # Drop rows with missing weight. Record drop
      n_weight_dropped <- 0L
      if (has_weight) {
        na_wt <- is.na(model_df[['weight_value']])
        n_weight_dropped <- sum(na_wt)
        if (n_weight_dropped > 0) {
          cat('\nDropping', n_weight_dropped, 'row(s) with NA weight_value')
          model_df <- model_df[!na_wt, , drop = FALSE]
        }
      }

      # Check if there are too few values or years
      n_years <- dplyr::n_distinct(model_df$year)
      n_vals  <- dplyr::n_distinct(model_df[[var]])
      if (n_years < 2 || n_vals < 2) {
        msg <- paste0(
          'Issue with ', var, ': ', n_years, ' distinct year(s), ',
          n_vals, ' distinct value(s) reaching the model'
        )
        warning(msg)
        return(msg)
      }

      # Report if state_code is NA. It shouldn't be
      if (anyNA(model_df$state_code)) {
        bad <- unique(model_df$fips[is.na(model_df$state_code)])
        msg <- paste0(
          'NA state_code ', var, ': ', length(bad),
          ' fips unmatched (', paste(head(bad), collapse = ', '), ')'
        )
        warning(msg)
        return(msg)
      }

      fp <- model_df$fips

      # Stop if fips code not in fp
      stopifnot(all(fp %in% rownames(W_county)))
      W_stack <- W_county[fp, fp]

      # No cross-year links
      W_stack[outer(model_df$year, model_df$year, `!=`)] <- 0
      diag(W_stack) <- 0
      lw <- mat2listw(W_stack, style = 'W', zero.policy = TRUE)
      cat('\nCreated lw')

      # Assign weights arg if variable is weighted. Otherwise NULL
      if (has_weight) {
        weights_arg <- model_df[['weight_value']]
      } else {
        weights_arg <- NULL
      }

      # Run error sar model with appropriate (or no) weighting
      model <- errorsarlm(
        normed ~ as.numeric(year) + state_code,
        data = model_df,
        listw = lw,
        weights = weights_arg,
        zero.policy = TRUE
      )
      cat('\nRan model')

      # Carry the NA-weight drop count with the model for reporting
      attr(model, 'n_weight_dropped') <- n_weight_dropped
      return(model)
      },
      error = function(e) {
        return(paste0('Error with ', var, ': ', e))
      }
    )
  }, .progress = FALSE, .options = furrr_options(seed = TRUE)) %>%
    setNames(vars)

  plan(sequential)
  toc()

  saveRDS(sar_outs, sar_path)
}
# 5 minutes or so
```

#### Reactable Table

First we wrangle our outputs to get the relevant results in a nicely formatted table:

```{r}
#| output: false
# Reload so this chunk runs when interactive re-run
sar_outs <- readRDS(sar_path)
(start <- length(sar_outs))
sar_keeps <- sar_outs[names(sar_outs) != 'state_code'] %>%
  keep(~ inherits(.x, 'Sarlm'))
(end <- length(sar_keeps))
(lost <- start - end)
# 42 to 37

# Which are missing and why
(lost_vars <- setdiff(names(sar_outs), names(sar_keeps)))
(reasons <- sar_outs[lost_vars] %>% keep(is.character))
# Losing 5 that are actually single year

# Weight rows dropped per model
(weight_drops <- imap(sar_keeps, ~ {
  n <- attr(.x, 'n_weight_dropped')
  tibble(variable_name = .y, n_weight_rows_dropped = if (is.null(n)) 0L else n)
}) %>%
  bind_rows() %>%
  filter(n_weight_rows_dropped > 0) %>%
  arrange(desc(n_weight_rows_dropped)))
# These had issues with missing weights - just removed those rows

sar_df <- imap(sar_keeps, ~ {
  tryCatch(
    {
      df <- tidy(.x)
      est <- df[2, 'estimate'] %>% as.numeric
      se <- df[2, 'std.error'] %>% as.numeric
      alpha <- 0.05
      z <- qnorm(1 - alpha/2)
      confints <- c(est - z * se, est + z * se)
      
      # confints <- confint(.x)[3, ]
      out <- df %>% 
        filter(str_detect(term, 'year')) %>% 
        mutate(
          variable_name = .y,
          conf_low = confints[1],
          conf_high = confints[2]
        ) %>% 
        select(-term) %>% 
        mutate(
          sig = case_when(
            p.value < 0.001 ~ '***',
            p.value > 0.001 & p.value < 0.01 ~ '**',
            p.value > 0.01 & p.value < 0.05 ~ '*',
            p.value > 0.05 ~ ' ',
            .default = NA
          )
        )
      return(out)
    },
      error = function(e) {
        message('Error')
        print(e)
    }
  )
}) %>% 
  keep(is.data.frame) %>% 
  bind_rows()
str(sar_df)
sar_df

(lost_on_wrangle <- setdiff(names(sar_keeps), sar_df$variable_name))
(check <- sar_outs[names(sar_keeps) %in% lost_on_wrangle])
# No more lost

sar_raw_outputs <- sar_df %>% 
  mutate(trend = case_when(
    estimate > 0 & str_detect(sig, '\\*') ~ 'increasing',
    estimate < 0 & str_detect(sig, '\\*') ~ 'decreasing',
    str_detect(sig, '\\*', negate = TRUE) ~ 'no change',
    .default = NA_character_
  ))
str(sar_raw_outputs)

# Bring in directional values. Keep only the ones without question marks, 
# clearly higher or lower
str(dp_meta)
directions <- dp_meta %>% 
  filter(str_detect(desirable, 'higher|lower'), !is.na(variable_name)) %>% 
  mutate(desirable = str_remove_all(desirable, '\\?')) %>% 
  select(variable_name, desirable)
reverse <- directions %>% 
  filter(desirable == 'lower') %>% 
  pull(variable_name)

# Reduce to only vars that have a direction.
# Here we lose a few metrics
sar_model_outputs <- sar_raw_outputs %>% 
  filter(variable_name %in% directions$variable_name) %>% 
  mutate(
    reverse = ifelse(variable_name %in% reverse, TRUE, FALSE),
    outcome = case_when(
      reverse == FALSE & trend == 'increasing' ~ 'better',
      reverse == FALSE & trend == 'decreasing' ~ 'worse',
      reverse == TRUE & trend == 'increasing' ~ 'worse',
      reverse == TRUE & trend == 'decreasing' ~ 'better',
      trend == 'no change' ~ 'no change',
      .default = NA
    ),
    across(where(is.numeric), ~ format(round(.x, 3), nsmall = 3))
  )
sar_model_outputs
nrow(sar_model_outputs) # 33

(n_metrics <- dim(sar_model_outputs)[1])

# See what we lost
(lost <- setdiff(sar_raw_outputs$variable_name, sar_model_outputs$variable_name))
(n_lost <- length(lost))
(print_lost <- lost %>% paste0(collapse = ', '))
# lost 4 without direction for better or worse

# proportion getting better or worse
(tab <- table(sar_model_outputs$outcome))
(percs <- round(prop.table(tab), 3) * 100)
```

We have `{r} n_metrics` metrics modeled. `{r} percs[['better']]`% are getting better, `{r} percs[['worse']]`% are getting worse, and `{r} percs[3]`% are stable. We drop `{r} length(lost)` metrics here that don't have proper directions (`{r} print_lost`).

Now we present the table, including normative outcomes:

```{r}
#| label: errorsar_reactable
sar_framed_outputs <- dp_meta %>% 
  select(dimension, index, indicator, metric, variable_name) %>% 
  right_join(sar_model_outputs, by = join_by(variable_name == variable_name))

## Create interactive table for Quarto
get_reactable(
  sar_framed_outputs, 
  defaultPageSize = 10,
  columns = list(
   variable_name = colDef(minWidth = 200),
   sig = colDef(minWidth = 50),
   reverse = colDef(minWidth = 75),
   statistic = colDef(minWidth = 75),
   outcome = colDef(minWidth = 75)
  )
)
```


#### LaTeX Table

Saving table to $\LaTeX$ for paper:

```{r}
#| label: latex_table
#| output: false
str(sar_framed_outputs)

# Reduce to specific columns
latex_df <- sar_framed_outputs %>% 
  select(metric, estimate:p.value, sig, outcome) %>% 
  mutate(
    p.value = case_when(
      str_detect(sig, '\\*') ~ paste0(p.value, sig),
      .default = p.value
    ),
    outcome = case_when(
      outcome == 'no change' ~ '\\textemdash',
      .default = str_to_title(outcome)
    )
  ) %>% 
  select(-sig) %>% 
  setNames(c(
    'Metric',
    '$\\beta$',
    'SE',
    '$z$-value',
    'p-value',
    'Outcome'
  ))
str(latex_df)

# Save to latex
latex_df %>% 
  kbl(
    format = 'latex',
    caption = 'Spatial error models of year and state on metric',
    label = 'tab_errorsarlm',
    booktabs = TRUE,
    escape = FALSE
  ) %>% 
  kable_styling(
    font_size = 9
  ) %>% 
  footnote(
    general_title = "",
    general = 
      '\\\\textit{Note: }The table shows coefficients of the effect of year on
      normalized, inflation-adjusted metrics with state as a covariate. 
      Metrics are weighted by variables shown in Table \\\\ref{tab:tab_metrics_body}. 
      Significant p-values
      indicated change over time, and the direction of the outcome was determined
      based on desirable directions of change for each metric. ***: p < 0.001,
      **: p < 0.01, *: p < 0.05.',
    threeparttable = TRUE,
    escape = FALSE
  ) %>% 
  save_kable(
    file = '4_trends/output/tab_errorsarlm.tex'
  )
```

#### Figure

Finally, the figure below shows coefficient estimates and confidence intervals for the errorsar models. Green points represent metrics that are trending in a "good" direction, while purple represents trends in a "bad" direction.

```{r}
#| label: errorsar_figure
# Using functions at top of script to process outputs and make graph
errorsar_figure_data <- prepare_figure_data(model_output_df = sar_framed_outputs)
errorsar_figure <- pct_change_graph(
  figure_data = errorsar_figure_data$figure_data,
  dp_text_palette = dp_text_palette,
  label_dimension_lookup = errorsar_figure_data$label_dimension_lookup
)

ggsave(
  '4_trends/output/fig_errorsarlm.png',
  plot = errorsar_figure,
  width = 8.5,
  height = 6,
  units = 'in',
  bg = 'white'
)

errorsar_figure
```
