---
title: "Metric Analysis"
format:
  html:
    fig-dpi: 200
    code-fold: true
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: false
---

```{r}
#| label: prep
#| cache: false
#| output: false
options(scipen = 999)

pacman::p_load(
  conflicted,
  dplyr,
  htmltools,
  stringr,
  tidyr,
  reactable,
  purrr,
  broom,
  knitr,
  kableExtra,
  tictoc,
  furrr,
  parallelly,
  splm,
  spdep,
  spatialreg,
  sf,
  skimr,
  ggplot2,
  ggpubr,
  readr,
  lmtest,
  sandwich,
  performance,
  snakecase,
  ggtext
)

pacman::p_load_current_gh(
  'ChrisDonovan307/projecter', 
  'Food-Systems-Research-Institute/SMdata'
)

# Load SMdocs project
devtools::load_all()

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::pull(),
  stats::lag(),
  base::setdiff(),
  .quiet = TRUE
)
```

```{r}
#| label: graph_functions
#| output: false
#| echo: false

# Take formatted figure data, a palette for text colors, the label dimension
# lookup from prepare_figure_data(). Return ggplot graph object
pct_change_graph <- function(figure_data, 
                             dp_text_palette, 
                             label_dimension_lookup) {

  # Renaming input to preserve it for error checking
  data <- figure_data
  
  # Standardize column names
  if ("conf_low" %in% names(data) && "conf_high" %in% names(data)) {
    data <- data %>%
      rename(ci_low = conf_low, ci_high = conf_high)
  }

  data <- data %>%
    mutate(across(c(estimate, ci_low, ci_high), ~ as.numeric(.x))) %>%
    mutate(
      across(c(estimate, starts_with('ci|conf')), as.numeric),
      outcome = ifelse(is.na(outcome), 'no change', outcome),
      outcome = snakecase::to_title_case(outcome)
    )

  ggplot(data, aes(x = estimate, y = combined_label)) +
    geom_point(aes(color = outcome)) +
    geom_errorbarh(aes(xmin = ci_low, xmax = ci_high, color = outcome)) +
    geom_vline(xintercept = 0) +
    
    # Invisible points to create a second legend (dimension colors)
    geom_point(
      aes(fill = dimension, color = dimension), 
      shape = 22, 
      alpha = 0, 
      show.legend = TRUE
    ) +
    
    # Add stars
    geom_text(
      aes(x = ci_high + 0.1, label = sig),
      hjust = 0,
      size = 4
    ) +
    
    # Outcome legend (color of real points and lines)
    scale_color_manual(
      values = c('Better' = 'darkgreen', 'No Change' = 'grey', 'Worse' = 'purple'),
      name = "Outcome"
    ) +
    
    # Dimension legend (fill colors)
    scale_fill_manual(
      values = dp_text_palette,
      name = "Dimension",
      labels = snakecase::to_title_case(names(dp_text_palette))
    ) +
    
    # Add both legends
    guides(
      color = guide_legend(order = 1, nrow = 2),
      fill = guide_legend(
        override.aes = list(shape = 22, size = 4, alpha = 1, color = dp_text_palette),
        order = 2,
        nrow = 2
      )
    ) +
    
    theme_minimal() +
    
    labs(
      x = 'Average Percent Change per Year',
      y = 'Indicator with Metric'
    ) + 
    
    scale_y_discrete(
      labels = make_y_axis_labels(label_dimension_lookup, dp_text_palette)
    ) +
    
    theme(
      legend.position = "top",
      legend.direction = "horizontal",
      legend.title.position = "top",
      text = element_text(family = "Times New Roman"),
      axis.title = element_text(family = "Times New Roman"),
      axis.text = element_text(family = "Times New Roman"),
      legend.title = element_text(family = "Times New Roman", hjust = 0.5),
      legend.text = element_text(family = "Times New Roman"),
      axis.text.y.left = ggtext::element_markdown(),
    )
}
```

This page explores time series data at the state and county level, then runs some regressions to see whether metrics are improving or declining over time at the county level. We first run weighted OLS models with robust SEs in @sec-ols. Then in @sec-spdep, we use spatial error models weighted by both geography and weighting variables.

## Time Series

Plot time series of each metric that has at least two data points. These are unweighted values. LOESS curves are shown in blue, while individual county/state data are shown in black

### County Data

First, we prepare a vector names of our time series variables for counties:

```{r}
#| label: prep_series
#| output: false
# Prepare vector of county variables that are time series
get_str(SMdocs::dp_metrics_county_adj)
vars <- unique(SMdocs::dp_metrics_county_adj$variable_name)
series_vars <- map(vars, ~ {
  count <- SMdocs::dp_metrics_county_adj %>% 
    filter(variable_name == .x) %>% 
    pull(year) %>% 
    unique() %>% 
    length()
  if (count > 1) {
    return(.x)
  } else {
    return(NULL)
  }
}) %>% 
  keep(~ !is.null(.x)) %>%
  unlist()
usethis::use_data(series_vars, overwrite = TRUE)
```

Now we plot each graph individually and arrange them together:

```{r}
#| label: county_time_series
#| fig-height: 25
#| fig-width: 12
#| cache: true
plots <- map(series_vars, ~ {
  df <- SMdocs::dp_metrics_county_adj %>%
    filter(variable_name == .x) %>%
    ggplot(aes(x = year, y = value)) +
    geom_line(
      aes(group = fips),
      alpha = 0.3
    ) +
    geom_smooth(
      method = 'loess',
      lwd = 2
    ) +
    labs(
      x = 'Year',
      title = .x
    ) +
    theme_classic()
})

ggarrange(
  plotlist = plots,
  ncol = 4,
  nrow = 11
)
```

### State Data

Again, start by getting a vector of time series metrics for state data, then plot all metrics:

```{r}
#| label: state_time_series
#| fig-height: 12
#| fig-width: 12
vars <- unique(SMdocs::dp_metrics_state_adj$variable_name)
state_series_vars <- map(vars, ~ {
  count <- SMdocs::dp_metrics_state_adj %>% 
    filter(variable_name == .x) %>% 
    pull(year) %>% 
    unique() %>% 
    length()
  if (count > 1) {
    return(.x)
  } else {
    return(NULL)
  }
}) %>% 
  keep(~ !is.null(.x)) %>%
  unlist()
usethis::use_data(state_series_vars, overwrite = TRUE)

plots <- map(state_series_vars, ~ {
  df <- SMdocs::dp_metrics_state_adj %>%
    filter(variable_name == .x) %>%
    mutate(across(c(year, value), as.numeric)) %>% 
    ggplot(aes(x = year, y = value)) +
    geom_line(
      aes(group = fips),
      alpha = 0.3
    ) +
    geom_smooth(
      method = 'loess',
      lwd = 2
      # col = '#154734'
    ) +
    labs(
      x = 'Year',
      title = .x
    ) +
    theme_classic()
})

ggarrange(
  plotlist = plots,
  ncol = 3,
  nrow = 5
)
```

### Tiny Trend Graphs

Outputs not shown here - these will be added to a table in the paper.

```{r}
all_series_vars <- c(series_vars, state_series_vars)

# Function to get tiny plots from county and state vars
get_tiny_plots <- function(df, vars) {
  map(vars, ~ {
    df <- df %>%
      filter(variable_name == .x, !is.na(value)) %>% 
      ggplot(aes(x = year, y = value)) +
      geom_line(
        aes(group = fips),
        alpha = 0.25,
        color = 'grey',
        linewidth = 0.5
      ) +
      geom_smooth(
        method = "loess", 
        se = FALSE, 
        linewdith = 0.75
      ) +
      coord_cartesian(xlim = c(2000, 2024)) +
      scale_x_discrete(expand = c(0, 0)) +
      scale_y_discrete(expand = c(0, 0)) +
      theme_bw() +
      theme(
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        plot.margin = margin(rep(0.5, 4))
        
      )
  }) %>% 
    setNames(c(vars))
}

# Plots for counties
county_plots <- SMdocs::dp_metrics_county_adj %>% 
  get_tiny_plots(
    vars = series_vars
  )

# Plots for states
state_plots <- SMdocs::dp_metrics_state_adj %>% 
  get_tiny_plots(
    vars = state_series_vars
  )

# One more plot for CPI - food at home
system_plots <- SMdata::metrics %>% 
  filter(variable_name == 'cpiFoodAtHome') %>% 
  get_tiny_plots(vars = 'cpiFoodAtHome')


# Save them all to outputs with variable names
iwalk(c(county_plots, state_plots, system_plots), \(plot, var) {
  ggsave(
    filename = paste0('../outputs/trend_plots/fig_trend_', var, '.png'),
    plot = plot,
    width = 1.5,
    height = 1.5,
    units = 'cm',
    dpi = 300,
    bg = NULL
  )
})
```

## Linear Models {#sec-ols}

First trying linear models with time as independent variable. Min max normalizing dataset with stored function before we do anything else.

```{r}
#| label: normalize
#| output: false
get_str(SMdocs::dp_metrics_county_adj)

vars <- unique(SMdocs::dp_metrics_county_adj$variable_name)
dp_metrics_county_adj_normed <- map(vars, ~ {
  SMdocs::dp_metrics_county_adj %>%
    filter(variable_name == .x) %>% 
    mutate(across(c(year, value), as.numeric)) %>% 
    drop_na(value) %>% 
    group_by(variable_name) %>% 
    mutate(normed_value = min_max(value)) %>% 
    ungroup()
}) %>% 
  list_rbind()
get_str(dp_metrics_county_adj_normed)  
usethis::use_data(dp_metrics_county_adj_normed, overwrite = TRUE)
```

### Regressions

First create a weights crosswalk that links `variable_name`s to weighting variables.

```{r}
#| label: weights_crosswalk
weights_crosswalk <- data.frame(
  metric = c(
    'GDP',
    'GDP from agriculture',
    'Total number of producers',
    'Total number of operations',
    'Population (annual)',
    'Population (5 year)',
    'Population (5 year, smooth)',
    'Acres of land',
    'Acres of water',
    'Area under agriculture (CDL)',
    'Number employed in agriculture',
    'Acres operated',
    'Acres operated, smooth',
    'Acres of cropland ag land',
    'Acres of woodland ag land',
    'Acres of pasture ag land',
    'Total farm income',
    'Total sales from commodities'
  ),
  variable_name = c(
    'gdp',
    'gdpFromAg',
    'nProducers',
    'nOperations',
    'populationAnnual',
    'population5Year',
    'population5YearSmooth',
    'landAcres',
    'waterAcres',
    'cdlAgAcres',
    'annualAvgEmplvlNAICS11',
    'acresOperated',
    'acresOperatedSmooth',
    'agLandCropAcres',
    'agLandWoodAcres',
    'agLandPastureAcres',
    'totalFarmIncome',
    'totalSalesCommodities'
  )
)

# Link weight vars to the appropriate variables
weights_crosswalk <- SMdocs::dp_meta %>% 
  select(variable_name, weighting) %>% 
  filter(variable_name %in% SMdocs::dp_metrics_county_adj_normed$variable_name) %>% 
  left_join(weights_crosswalk, by = join_by(weighting == metric)) %>% 
  setNames(c('variable_name', 'weight_metric', 'weight_variable')) %>% 
  select(-weight_metric) %>% 
  na.omit()
head(weights_crosswalk)
```

Run weighted OLS regressions for all time series variables where time points \> 1. The general formula is: 

$$
\Large y_i = \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i} + \epsilon_i,
$$

where $y_i$ is the normalized value of the metric across counties, $x_{1i}$ is the year, $x_{2i}$ is the state each county is part of, and $\epsilon$ is the error term.

```{r}
#| label: regressions
#| output: false
#| eval: false
# Get a crosswalk to add state as a fixed effect
state_link <- SMdata::fips_key %>% 
  filter(str_length(fips) == 2, fips != '00') %>% 
  select(fips, state_code)

res <- map(SMdocs::series_vars, ~ {
  tryCatch(
    {
      df <- SMdocs::dp_metrics_county_adj_normed %>% 
        filter(variable_name == .x) %>% 
        mutate(
          value = as.numeric(value),
          year = as.numeric(year),
          state_fips = str_sub(fips, end = 2)
        ) %>% 
        left_join(state_link, by = join_by(state_fips == fips))
      
      # Add weighting variable
      weight_var <- weights_crosswalk$weight_variable[weights_crosswalk$variable_name == .x]
      weight_df <- SMdocs::dp_weights
      if (!is.na(weight_var) && length(weight_var) > 0) {
        weight_df <- weight_df %>% 
          filter(variable_name == weight_var)
      }
      weight_df <- weight_df %>% 
        mutate(
          value = as.numeric(value), 
          year = as.numeric(year)
        ) %>% 
        rename(
          weight_value = value, 
          weight_name = variable_name,
          weight_year = year
        )
      # get_str(weight_df)
      
      # Depending on whether weight has a year, do appropriate join
      # Note that we might be losing some years here if weighting var doesn't
      # cover the variable enough
      if (all(is.na(weight_df$weight_year))) {
        df <- df %>% 
          inner_join(weight_df, by = join_by(fips))
      } else {
        df <- df %>% 
          inner_join(weight_df, by = join_by(fips, year == weight_year))
      }
      
      # Reduce to county fips only. Remove dupes 
      df <- df %>%
        filter(str_length(fips) == 5) %>% 
        distinct()
      
      # Weighted OLS 
      model <- lm(
        normed_value ~ year + state_code, 
        data = df,
        weights = weight_value
      )
      return(model)
    },
    error = function(e) {
      message(paste('Error with var', .x))
    }
  )
}) %>% 
  setNames(c(SMdocs::series_vars)) %>%
  discard(\(x) is.null(x))
get_str(res, 1)
get_str(res)

# Robust SEs
lmtest::coeftest(res[[1]])

# Save immediate output
saveRDS(res, '../data/objects/weighted_ols_outs.rds')
```

We are saving these outputs to a file so we don't have to rerun it every time we want them.

### Wrangling

Now that we have regression outputs, let's put them in a table to see which ones change significantly over time.

```{r}
#| label: make_reg_table
#| output: false
# Put together one table of results from all models, add stars
res <- readRDS('../data/objects/weighted_ols_outs.rds')
get_str(res)

# Check for NA outputs
res[[1]]$coefficients[['year']]
coefs <- map(res, ~ .x$coefficients[['year']])
names(coefs[is.na(coefs)])

# Mobility disability is NA because there is only 1 year after we join with 
# weighting variable. So we need to just throw it out if we can't get smoothed
# population for 2024.
res <- res %>% 
  discard(~ is.na(.x$coefficients[[2]]))

# Make a table of regression outputs, including robust standard errors
res_df <- imap(res, ~ {
  vcov = vcovHC(.x, type = 'HC0')
  df <- coeftest(.x, vcov = vcov) %>% 
    tidy() %>% 
    filter(term == 'year') %>% 
    mutate(term = .y)
  conf_ints <- coefci(.x)['year', ]
  df$ci_low <- conf_ints[1]
  df$ci_high <- conf_ints[2]
  df
}) %>% 
  bind_rows() %>% 
  mutate(sig = ifelse(p.value < 0.05, '*', ''))
res_df

# How many time series metrics total
(n_metrics <- nrow(res_df))

# How many change over time significantly
(
  n_change <- res_df %>% 
    filter(sig == '*') %>% 
    nrow()
)

# Percentage significant
(perc_sig <- mean(res_df$sig == '*', na.rm = TRUE) * 100)
```

Note that we only have 2 years for the mobility disability metric - 2023 and 2024. But we don't have a smoothed 5-year population for 2024, only 2023. So we can only use a single year, which means there is no variation in year. So we are dropping this from the analysis.

Out of `{r} n_metrics` metrics with \> 1 time point, we have `{r} n_change` metrics that change significantly over time (`{r} round(perc_sig, 2)`%).

Still have to make sense of the directional values of our metrics. These were defined in the [Aggregation](../pages/aggregation.qmd) from the main analysis branch. In this next chunk, we assign normative directions to each metric so we can tell whether they are getting better or worse over time based on whether high values or low values are desirable.

```{r}
#| label: make_sense_of_trends
#| output: false
# Of those that are significant, what are trends
# leaving in all of them, even if not significant though
raw_outputs <- res_df %>% 
  mutate(trend = case_when(
    estimate > 0 & sig == '*' ~ 'increasing',
    sig == '' ~ NA_character_,
    .default = 'decreasing'
  ))

# Bring in directional values. Keep only the ones without question marks, 
# clearly higher or lower
get_str(SMdocs::dp_meta)
directions <- SMdocs::dp_meta %>% 
  filter(str_detect(desirable, 'higher|lower'), !is.na(variable_name)) %>% 
  mutate(desirable = str_remove_all(desirable, '\\?')) %>% 
  select(variable_name, desirable)
reverse <- directions %>% 
  filter(desirable == 'lower') %>% 
  pull(variable_name)

# Reduce to only vars that have a direction.
# Here we lose a few metrics
model_outputs <- raw_outputs %>% 
  filter(term %in% directions$variable_name) %>% 
  mutate(
    reverse = ifelse(term %in% reverse, TRUE, FALSE),
    outcome = case_when(
      reverse == FALSE & trend == 'increasing' ~ 'better',
      reverse == FALSE & trend == 'decreasing' ~ 'worse',
      reverse == TRUE & trend == 'increasing' ~ 'worse',
      reverse == TRUE & trend == 'decreasing' ~ 'better',
      .default = NA
    ),
    across(where(is.numeric), ~ format(round(.x, 3), nsmall = 3))
  )
model_outputs

# See what we lost
(lost <- setdiff(raw_outputs$term, model_outputs$term))
(n_lost <- length(lost))
(print_lost <- lost %>% paste0(collapse = ', '))

# proportion getting better or worse
(tab <- table(model_outputs$outcome))
(percs <- round(prop.table(tab), 3) * 100)
```

`{r} percs[['better']]`% are getting better and `{r} percs[['worse']]`% are getting worse. We drop `{r} length(lost)` metrics here that don't have proper directions (`{r} print_lost`).


### Table

Now we can check out the outputs table for each regression. In each model, 


Trend is just whether it is going up or down, outcome is whether that is good or bad.

```{r}
#| label: reg_table
# Add framework info
framed_outputs <- SMdocs::dp_meta %>% 
  select(dimension, index, indicator, metric, variable_name) %>% 
  right_join(model_outputs, by = join_by(variable_name == term))

# Create interactive table for Quarto
get_reactable(
  framed_outputs, 
  defaultPageSize = 10,
  columns = list(
   variable_name = colDef(minWidth = 200),
   sig = colDef(minWidth = 50),
   reverse = colDef(minWidth = 75),
   statistic = colDef(minWidth = 75),
   outcome = colDef(minWidth = 75)
  )
)
```

### Figure

Figure below shows 95% confidence intervals for the coefficients of each metric, including state as a covariate.

```{r}
#| label: regression_figure
# Combined label for indicator and metric

# See repository for function definitions!
processed_lm_data <- prepare_figure_data(framed_outputs)

lm_figure <- pct_change_graph(
  figure_data = processed_lm_data$figure_data,
  dp_text_palette = dp_text_palette,
  label_dimension_lookup = processed_lm_data$label_dimension_lookup
)

ggsave(
  '../outputs/fig_lm.png',
  plot = lm_figure,
  width = 8,
  height = 5,
  units = 'in',
  bg = 'white'
)

lm_figure
```

## Spatial Regressions {#sec-spdep}

### Models

Here we run the same process but using spatial error models with the `spatialreg` package. We are now working with:

$$
\Large y = X\beta + u, u = \lambda Wu + \epsilon,
$$

where $y$ are normalized metric values, $X$ is an $n * k$ matrix, $\beta$ is a vector of predictors, $u$ is the spatial autocorrelation error term, $\lambda$ is the spatial error parameter, $W$ is the county weights matrix, and $\epsilon$ is the error term. 

We run this portion of the code in parallel because it has a pretty long run time. Again, results are saved so it doesn't need to be run again unless something changes.

```{r}
#| label: spatial_reg_all
#| eval: false
#| output: false
get_str(SMdocs::dp_metrics_county_adj)
state_link <- SMdata::fips_key %>% 
  filter(str_length(fips) == 2, fips != '00') %>% 
  select(fips, state_code)

dat <- SMdocs::dp_metrics_county_adj %>% 
  pivot_wider(
    id_cols = fips:year,
    names_from = variable_name,
    values_from = value
  ) %>% 
  filter(str_detect(fips, '^09', negate = TRUE)) %>% 
  mutate(link = str_sub(fips, end = 2)) %>% 
  left_join(state_link, by = join_by(link == fips)) %>% 
  select(-link) %>% 
  mutate(across(!c(fips, state_code), as.numeric))
get_str(dat)

# Get vector of vars
(vars <- names(dat)[!names(dat) %in% c('fips', 'year', 'state_code')])

tic()
get_time()
plan(multisession, workers = availableCores(omit = 2))
# out <- map(vars[1], \(var) {
out <- future_map(vars, \(var) {
  cat('\n\nStarting', var)
  
  tryCatch({
    # Reduce to relevant variable, remove NAs
    df <- dat %>% 
      select(fips, year, !!sym(var), state_code) %>% 
      drop_na(!!sym(var)) %>% 
      mutate(normed = min_max(!!sym(var)))
 
    # Regular weights
    weight_var <- weights_crosswalk$weight_variable[weights_crosswalk$variable_name == var]
    if (!is.na(weight_var) && length(weight_var) > 0) {
      weight_df <- SMdocs::dp_weights %>% 
        mutate(
          value = as.numeric(value), 
          year = as.numeric(year)
        ) %>% 
        rename(
          weight_value = value, 
          weight_name = variable_name,
          weight_year = year
        ) %>% 
        filter(weight_name == weight_var, str_length(fips) == 5)
      
      # Join weights back to df
      # df <- df %>% 
      #   inner_join(weight_df, by = join_by(fips, year == weight_year))
      if (all(is.na(weight_df$weight_year))) {
        df <- df %>% 
          inner_join(weight_df, by = join_by(fips))
      } else {
        df <- df %>% 
          inner_join(weight_df, by = join_by(fips, year == weight_year))
      }
      
      cat('\nJoined to weights:', weight_var)
    } else {
      cat('\nNo weighting variable')
    }
    
    # List weights
    lw <- SMdata::neast_counties_2024 %>% 
      right_join(df) %>% 
      select(fips, year, normed) %>% 
      poly2nb() %>% 
      nb2listw(zero.policy = TRUE)   
    cat('\nCreated lw')
    
    # Assign weights arg if variable is weighted. Otherwise NULL
    if (!is.na(weight_var) && length(weight_var) > 0) {
      weights_arg <- df[['weight_value']]
    } else {
      weights_arg <- NULL
    }
    
    # Run error sar model with appropriate (or no) weighting
    model <- errorsarlm(
      normed ~ as.numeric(year) + state_code,
      data = df,
      listw = lw,
      weights = weights_arg
    ) 
    cat('\nRan model')
   
    return(model)
    },
    error = function(e) {
      return(paste0('Error with ', var, ': ', e))
    }
  )
}, .progress = FALSE) %>% 
  setNames(c(vars))
plan(sequential)
toc()
# 6 to 12 minutes or so

# out
# out[[1]]

get_str(out)
get_str(out[[1]])
map(out, class)
# get_str(out[[1]])
# map(out, ~ .x['fips'])
# map(out, ~ summary(.x))

saveRDS(out, '../data/objects/errorsarlm_outs.rds')
```

#### Reactable Table

First we wrangle our outputs to get the relevant results in a nicely formatted table:

```{r}
#| output: false
sar_outs <- readRDS('../data/objects/errorsarlm_outs.rds')
start <- length(sar_outs)
sar_outs <- sar_outs[names(sar_outs) != 'state_code'] %>% 
  keep(~ class(.x) == 'Sarlm')
end <- length(sar_outs)
(lost <- start - end)
# 42 down to 38 here

sar_df <- imap(sar_outs, ~ {
  tryCatch(
    {
      df <- tidy(.x)
      est <- df[2, 'estimate'] %>% as.numeric
      se <- df[2, 'std.error'] %>% as.numeric
      alpha <- 0.05
      z <- qnorm(1 - alpha/2)
      confints <- c(est - z * se, est + z * se)
      
      # confints <- confint(.x)[3, ]
      out <- df %>% 
        filter(str_detect(term, 'year')) %>% 
        mutate(
          variable_name = .y,
          conf_low = confints[1],
          conf_high = confints[2]
        ) %>% 
        select(-term) %>% 
        mutate(
          sig = case_when(
            p.value < 0.001 ~ '***',
            p.value > 0.001 & p.value < 0.01 ~ '**',
            p.value > 0.01 & p.value < 0.05 ~ '*',
            p.value > 0.05 ~ ' ',
            .default = NA
          )
        )
      return(out)
    },
      error = function(e) {
        message('Error')
        print(e)
    }
  )
}) %>% 
  keep(is.data.frame) %>% 
  bind_rows()
get_str(sar_df)
# sar_df

# 38 down to 36.
(lost <- setdiff(names(sar_outs), sar_df$variable_name))
check <- sar_outs[names(sar_outs) %in% lost]
# Not defined because of singularities. Single time points. 
# At least while including the weighting variables.

sar_raw_outputs <- sar_df %>% 
  mutate(trend = case_when(
    estimate > 0 & str_detect(sig, '\\*') ~ 'increasing',
    estimate < 0 & str_detect(sig, '\\*') ~ 'decreasing',
    str_detect(sig, '\\*', negate = TRUE) ~ 'no change',
    .default = NA_character_
  ))
get_str(sar_raw_outputs)

# Bring in directional values. Keep only the ones without question marks, 
# clearly higher or lower
get_str(SMdocs::dp_meta)
directions <- SMdocs::dp_meta %>% 
  filter(str_detect(desirable, 'higher|lower'), !is.na(variable_name)) %>% 
  mutate(desirable = str_remove_all(desirable, '\\?')) %>% 
  select(variable_name, desirable)
reverse <- directions %>% 
  filter(desirable == 'lower') %>% 
  pull(variable_name)

# Reduce to only vars that have a direction.
# Here we lose a few metrics
sar_model_outputs <- sar_raw_outputs %>% 
  filter(variable_name %in% directions$variable_name) %>% 
  mutate(
    reverse = ifelse(variable_name %in% reverse, TRUE, FALSE),
    outcome = case_when(
      reverse == FALSE & trend == 'increasing' ~ 'better',
      reverse == FALSE & trend == 'decreasing' ~ 'worse',
      reverse == TRUE & trend == 'increasing' ~ 'worse',
      reverse == TRUE & trend == 'decreasing' ~ 'better',
      trend == 'no change' ~ 'no change',
      .default = NA
    ),
    across(where(is.numeric), ~ format(round(.x, 3), nsmall = 3))
  )
sar_model_outputs

(n_metrics <- dim(sar_model_outputs)[1])

# See what we lost
(lost <- setdiff(sar_raw_outputs$variable_name, sar_model_outputs$variable_name))
(n_lost <- length(lost))
(print_lost <- lost %>% paste0(collapse = ', '))

# proportion getting better or worse
(tab <- get_table(sar_model_outputs$outcome))
(percs <- round(prop.table(tab), 3) * 100)
```

We have `{r} n_metrics` metrics modeled. `{r} percs[['better']]`% are getting better, `{r} percs[['worse']]`% are getting worse, and `{r} percs[3]`% are stable. We drop `{r} length(lost)` metrics here that don't have proper directions (`{r} print_lost`).

Now we present the table, including normative outcomes:

```{r}
#| label: errorsar_reactable
sar_framed_outputs <- SMdocs::dp_meta %>% 
  select(dimension, index, indicator, metric, variable_name) %>% 
  right_join(sar_model_outputs, by = join_by(variable_name == variable_name))

## Create interactive table for Quarto
get_reactable(
  sar_framed_outputs, 
  defaultPageSize = 10,
  columns = list(
   variable_name = colDef(minWidth = 200),
   sig = colDef(minWidth = 50),
   reverse = colDef(minWidth = 75),
   statistic = colDef(minWidth = 75),
   outcome = colDef(minWidth = 75)
  )
)
```


#### LaTeX Table

Saving table to $\LaTeX$ for paper:

```{r}
#| label: latex_table
#| output: false
get_str(sar_framed_outputs)

# Reduce to specific columns
latex_df <- sar_framed_outputs %>% 
  select(metric, estimate:p.value, sig, outcome) %>% 
  mutate(
    # p.value = ifelse(sig == '*', paste0(p.value, '*'), p.value),
    p.value = case_when(
      str_detect(sig, '\\*') ~ paste0(p.value, sig),
      .default = p.value
    ),
    outcome = case_when(
      outcome == 'no change' ~ '\\textemdash',
      .default = str_to_title(outcome)
    )
  ) %>% 
  select(-sig) %>% 
  setNames(c(
    'Metric',
    '$\\beta$',
    'SE',
    '$z$-value',
    'p-value',
    'Outcome'
  ))
get_str(latex_df)

# Save to latex
latex_df %>% 
  kbl(
    format = 'latex',
    caption = 'Spatial error models of year and state on metric',
    label = 'tab_errorsarlm',
    booktabs = TRUE,
    escape = FALSE
  ) %>% 
  kable_styling(
    font_size = 9
  ) %>% 
  footnote(
    general_title = "",
    general = 
      '\\\\textit{Note: }The table shows coefficients of the effect of year on
      normalized, inflation-adjusted metrics with state as a covariate. 
      Metrics are weighted by variables shown in Table \\\\ref{tab:tab_metrics_body}. 
      Significant p-values
      indicated change over time, and the direction of the outcome was determined
      based on desirable directions of change for each metric. ***: p < 0.001,
      **: p < 0.01, *: p < 0.05.',
    threeparttable = TRUE,
    escape = FALSE
  ) %>% 
  save_kable(
    file = '../outputs/tab_errorsarlm.tex'
  )
```

#### Figure

Finally, the figure below shows coefficient estimates and confidence intervals for the errorsar models. Green points represent metrics that are trending in a "good" direction, while purple represents trends in a "bad" direction.

```{r}
#| label: errorsar_figure
# Using functions at top of script to process outputs and make graph
errorsar_figure_data <- prepare_figure_data(model_output_df = sar_framed_outputs)
errorsar_figure <- pct_change_graph(
  figure_data = errorsar_figure_data$figure_data,
  dp_text_palette = dp_text_palette,
  label_dimension_lookup = errorsar_figure_data$label_dimension_lookup
)

ggsave(
  '../outputs/fig_errorsarlm.png',
  plot = errorsar_figure,
  width = 8.5,
  height = 6,
  units = 'in',
  bg = 'white'
)

errorsar_figure
```

