---
title: "Coverage"
format:
  html:
    fig-dpi: 200
    code-fold: true
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: false
---

```{r}
#| label: prep
#| cache: false
#| output: false
options(scipen = 999)

pacman::p_load(
  dplyr,
  htmltools,
  stringr,
  tidyr,
  purrr,
  broom,
  ggplot2,
  ggpubr,
  tibble,
  Hmisc,
  reshape2,
  viridisLite,
  conflicted,
  readr,
  sf,
  tmap,
  mapview,
  snakecase,
  knitr,
  kableExtra
)

# GitHub packages
pacman::p_load_current_gh(
  'Food-Systems-Research-Institute/SMdata',
  'Food-Systems-Research-Institute/SMdocs'
)

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::pull(),
  dplyr::summarize(),
  stats::lag(),
  base::setdiff(),
  .quiet = TRUE
)
```

Here we explore how secondary data sources cover the region by dimension, geography, and over time.


## Dimensional

Explore coverage by dimension. Things we are interested in:

- Proportion of indicators in each dimension that have at least one metric
- Same for county metric vs state
- Which indicators have no metrics

Indicators without any metrics:

```{r}
SMdocs::dp_meta %>% 
  filter(is.na(variable_name)) %>% 
  select(dimension, indicator) %>% 
  unique() %>% 
  get_reactable()
```

### Representation and Friends

Proportion of indicator coverage by dimension:

```{r}
#| label: indicator_coverage_by_dimension
# Dimensions, indicators, metrics, and indicator representation
tab <- SMdocs::dp_meta %>% 
  group_by(dimension, indicator) %>% 
  summarize(
    has_metric = any(!is.na(variable_name)),
    n_metrics = sum(metric != 'NONE'),
    .groups = 'drop_last'
  ) %>% 
  summarize(
    indicators = n(),
    perc_rep_ind = sum(has_metric) / indicators * 100,
    metrics = sum(n_metrics),
    rep_ind = sum(has_metric),
    .groups = "drop"
  ) 
tab

# County metrics
county_metrics <- SMdocs::dp_metrics_county %>% 
  select(variable_name) %>% 
  distinct()
dim(county_metrics)
county_col <- SMdocs::dp_meta %>% 
  select(dimension, variable_name) %>% 
  right_join(county_metrics) %>% 
  group_by(dimension) %>% 
  summarize(`County Scale` = n())
tab <- tab %>% 
  inner_join(county_col)

# Add row of totals
totals <- tab %>% 
  summarize(across(c(indicators:metrics, `County Scale`), base::sum))
totals

str(tab)
latex_tab <- tab %>% 
  add_row(
    'dimension' = 'Total',
    'indicators' = as.double(totals[1, 'indicators']),
    'metrics' = as.double(totals[1, 'metrics']),
    'perc_rep_ind' = sum(.[['rep_ind']]) / sum(.[['indicators']]) * 100,
    'County Scale' = as.double(totals[1, 'County Scale'])
  ) %>% 
  select(-rep_ind) %>% 
  setNames(c(
    names(.) %>% 
      str_to_title() %>% 
      str_replace('Perc_rep_ind', 'Representation (\\\\%)')
  ))
latex_tab

# Save to latex
latex_tab %>% 
  kbl(
    format = 'latex',
    caption = 'Indicators and metrics by dimension.',
    label = 'tab_metric_breakdown',
    booktabs = TRUE,
    escape = FALSE,
    digits = 1
  ) %>% 
  kable_styling(
    font_size = 10
  ) %>% 
  footnote(
    general_title = "",
    general = 
      '\\\\textit{Note:} Representation is the percentage of indicators that are represented by at least one metric in the available data. The County Scale column shows metrics that are available at or below the county level and included in correlation and regression analyses.',
    threeparttable = TRUE,
    escape = FALSE
  ) %>% 
  row_spec(6, bold = TRUE) %>% 
  save_kable(
    file = '3_coverage/output/tab_metric_breakdown.tex'
  )

# Show table
tab %>% 
  get_reactable()
```

Looks like all Economics and Health indicators had a secondary metric to represent them. Environment and Production were missing a few, and Social is missing 60%.

How many metrics are at the state level vs county level? Note that we are prioritizing county and only counting state if they are not available at the county level.

```{r}
dat <- SMdocs::dp_meta %>% 
  filter(
    !is.na(variable_name),
    metric != 'NONE'
  ) %>% 
  select(
    dimension,
    metric,
    resolution
  ) %>% 
  mutate(resolution = case_when(
    resolution %in% c('state', 'system') ~ 'state',
    .default = 'county'
  ))

# Overall
overall_county <- round(mean(dat$resolution == 'county') * 100, 2)
overall_state <- 100 - overall_county
```

Overall, `{r} overall_county`% of metrics were at the county level, and `{r} overall_state`% were at the state level.

Which ones were state level:

```{r}
get_reactable(dat)
```

Break down by dimension:

```{r}
# By Dimension:
dat %>% 
  group_by(dimension) %>% 
  summarize(
    prop_county = mean(resolution == 'county')
    # prop_state = mean(resolution == 'state')
  ) %>% 
  mutate(across(starts_with('prop'), ~ format(round(.x, 3), nsmall = 3))) %>% 
  get_reactable()
```

We can see that the Economics, Health, and Social dimensions have rather complete data across counties, but Social and Production are missing a substantial amount.

## Geographic

### Summary Stats

Things we are interested in here:

- How many metrics were represented in each county? Breakdown by dimension
- Which counties have least metrics
- Which metrics are only at state level

First wrangle up a Start with metrics in each county, including a population county to explore:

```{r}
#| label: geo_summary_stats
# str(SMdocs::dp_metrics_county)
n_county_metrics <- SMdocs::dp_metrics_county$variable_name %>% 
  unique %>% 
  length

# Latest population of each county
pops <- SMdocs::dp_weights %>% 
  filter(variable_name == 'population5YearSmooth', year == '2023') %>% 
  select(fips, value)

# Metrics per county
# str(SMdocs::dp_metrics_county)
tab_metrics <- SMdocs::dp_metrics_county %>% 
  group_by(fips) %>% 
  summarize(n_metrics = length(unique(variable_name))) %>% 
  ungroup() %>% 
  left_join(SMdata::fips_key) %>% 
  select(-state_code) %>% 
  left_join(pops) %>% 
  rename(population_2023 = value) %>% 
  mutate(population_2023 = as.numeric(population_2023)) %>% 
  arrange(n_metrics)
get_reactable(tab_metrics)
```

Now break them down by dimension:

```{r}
dims <- SMdocs::dp_meta %>% 
  select(variable_name, dimension) %>% 
  na.omit() %>% 
  filter(variable_name != 'NONE')
tab_metrics_dim <- SMdocs::dp_metrics_county %>% 
  left_join(dims) %>% 
  group_by(fips, dimension) %>% 
  summarize(n_metrics = length(unique(variable_name)))
  # distinct(fips) %>%
  # left_join(pops, by = 'fips') %>% 
  # rename(population_2023 = value)
get_reactable(tab_metrics_dim)
```

Average number of metrics in each county in across years:

```{r}
# Get total number of county metrics
total_metrics <- SMdocs::dp_metrics %>% 
  filter(year >= 2000, year < 2025) %>% 
  SMdata::filter_fips('counties') %>% 
  pull(variable_name) %>% 
  unique() %>% 
  length()

# Metric counts by county and year
tab <- SMdocs::dp_metrics %>% 
  filter(year >= 2000, year < 2025) %>% 
  SMdata::filter_fips('counties') %>% 
  group_by(fips, year) %>% 
  summarize(
    n_metrics = length(unique(variable_name)),
    prop_metrics = n_metrics/total_metrics
  ) %>% 
  ungroup() %>% 
  left_join(fips_key) %>% 
  select(-state_code) %>% 
  arrange(desc(year)) %>%
  mutate(prop_metrics =  format(round(prop_metrics, 3), nsmall = 3)) %>% 
  filter(state_name != 'Connecticut')
get_reactable(tab)
```

Overall figures for metrics across years:

```{r}
# Average coverage of county overall
mean_metrics_overall <- round(mean(tab$n_metrics, na.rm = TRUE), 3)
prop_metrics_overall <- mean_metrics_overall/total_metrics
```

Average coverage of counties by state in any given year:

```{r}
tab %>% 
  group_by(state_name) %>% 
  summarize(
    mean_metrics = mean(n_metrics, na.rm = TRUE),
    prop_metrics = mean_metrics/total_metrics
  ) %>% 
  ungroup() %>% 
  arrange(mean_metrics) %>% 
  mutate(across(where(is.numeric), ~ format(round(.x, 3), nsmall = 3))) %>% 
  get_reactable()
```

Note on missing data: shouldn't be calculating it this way. 5-year data does not show up every year, even if it is complete. Need to rethink this. Expand grid to get total number of data points each county should have?

### Maps

Here we explore some maps to show missingness by geography. However, they end up not being terribly helpful due to the rather subtle patterns of missingness in our data. 

The first map shows the total number of metrics with data in each county across all five dimensions:

```{r}
#| output: false
## Get bounding box - commenting out to avoid downloading this each time
# would be better off caching though

# bbox_new <- st_bbox(neast_counties_2024)
# xrange <- bbox_new$xmax - bbox_new$xmin # range of x values
# yrange <- bbox_new$ymax - bbox_new$ymin # range of y values
# 
# bbox_new[1] <- bbox_new[1] - (0.05 * xrange) # xmin - left
# bbox_new[3] <- bbox_new[3] + (0.05 * xrange) # xmax - right
# bbox_new[2] <- bbox_new[2] - (0.05 * yrange) # ymin - bottom
# bbox_new[4] <- bbox_new[4] + (0.05 * yrange) # ymax - top
# 
# tiles <- get_tiles(
#   bbox_new,
#   provider = "CartoDB.PositronNoLabels",
#   zoom = 7,
#   crop = TRUE
# )
# terra::saveRDS(tiles, 'data/data_paper/neast_tiles.rds')

# Load tiles for background
tiles <- terra::readRDS('3_coverage/data/data_paper/neast_tiles.rds')

# Counts of unique metrics in each county
# var_counts <- datasets$dp_metrics_all %>% 
var_counts <- dp_metrics_county %>% 
  group_by(fips) %>% 
  summarize(n_metrics = length(unique(variable_name)))
 
# Join 
df <- neast_counties_2024 %>% 
  left_join(var_counts)

tmap_mode('plot')
map <- tm_shape(tiles) +
  tm_rgb() +
  tm_shape(df) +
  tm_polygons(
    "n_metrics", 
    palette = "brewer.greens",
    title = "Metric\nRepresentation",
    breaks = c(seq(22, 43, 3)),
    breaks = seq(min(var_counts$n_metrics), max(var_counts$n_metrics), 5),
    fill.legend = tm_legend(
      reverse = TRUE
    )
  ) +
  tm_borders(col = 'black', lwd = 1.25) +
  tm_layout(
    legend.position = c('left', 'top'),
    legend.title.fontface = 'bold',
    legend.width = 8,
    legend.height = 12,
    legend.title.size = 1.1,
    inner.margins = rep(0, 4),
    outer.margins = rep(0, 4),
    legend.text.size = 1
  )

# tmap_save(
#   tm = map,
#   filename = 'outputs/metric_coverage_map.png',
#   asp = 0,
#   dpi = 300
# )
```

```{r}
#| echo: false
#| fig-height: 8
#| fig-width: 8
#| out-width: '75%'
#| fig-align: 'center'
map
```

Connecticut falls out of this due to the county/region transition. For the rest of our counties, we don't really have the granularity to tell how different they are across dimensions. What we can see is that some of our most urban counties around New York and Boston are missing a handful of metrics.

Let's try an interactive map so we can explore each dimension individually. Use the layer symbol in the top left to select which dimension to view. These maps show the proportion of the total metrics available in each county.

```{r}
#| label: var_counts
#| output: false
# How many total metrics at county level
total_count <- dp_metrics_county %>% 
  pull(variable_name) %>% 
  unique %>% 
  length
total_count

# Dimension crosswalk
crosswalk <- dp_meta %>% 
  select(dimension, variable_name)

dimension_counts <- dp_meta %>% 
  select(dimension, variable_name, resolution) %>% 
  filter(!is.na(variable_name), resolution != 'state') %>% 
  select(-resolution) %>% 
  group_by(dimension) %>% 
  summarize(count = n())
dimension_counts
  
# Counts of unique metrics in each county by dimension
var_counts <- dp_metrics_county %>% 
  left_join(crosswalk) %>% 
  group_by(dimension, fips) %>% 
  summarize(
    n_metrics = length(unique(variable_name))
  ) %>% 
  left_join(dimension_counts) %>% 
  mutate(prop_metrics = n_metrics / count)
var_counts
 
# Join 
df <- neast_counties_2024 %>% 
  left_join(var_counts)
str(df)

maps <- df %>%
  split(.$dimension) %>%
  imap(~ {
    to_hide <- ifelse(.y == 'economics', FALSE, TRUE)
    mapview(
      .x, 
      layer.name = .y, 
      zcol = "prop_metrics",
      hide = to_hide,
      # col.regions = brewer.pal(5, "Greens"),
      col.regions = rev(viridis(5)),
      alpha.regions = 0.7
    )
  }) 
```

```{r}
#| echo: false
#| out-width: '90%'
#| fig-align: center
Reduce(`+`, maps)
```


## Temporal

This section explores the growth in metric and indicator counts over time. 

```{r}
#| include: false
# dat <- SMdocs::dp_metrics
# str(dat)
#   
# n_county_vars <- SMdocs::dp_metrics_county %>% 
#   pull(variable_name) %>% 
#   unique() %>% 
#   length()
# SMdocs::series_vars
# SMdocs::state_series_vars
```

### Indicators

Here we show the total number of indicators represented each year, colored by dimension. We can see the growth in represented indicators over time, as well as the 5-year update cycle of NASS and the American Community Survey. Credits to Isabella Loconte on this graph!

```{r}
#| label: Total indicators each year
#| out-width: '75%'
#| fig-align: center
# Join based on 'variable_name'
merged_df <- SMdocs::dp_metrics %>%
  left_join(SMdocs::dp_meta %>%
              select(variable_name,
                     dimension = dimension,
                     index = index,
                     indicator = indicator
              ),
            by = "variable_name") 

# Filter out before 2000 and NAs; create "indicator count"
indicators_by_year_data <- merged_df %>%
  filter(!is.na(indicator)) %>%
  filter(year >= 2000, year < 2025) %>%
  distinct(indicator, dimension, year) %>%
  count(year, dimension, name = "indicator_count") %>% 
  mutate(year = as.numeric(year))
str(indicators_by_year_data)

# Histogram
indicator_plot <- ggplot(
    indicators_by_year_data, 
    aes(x = year, y = indicator_count, fill = dimension)
  ) +
  geom_col(position = "stack") +
  scale_y_continuous(breaks = seq(0, max(1000), by = 5)) + 
  # max is set to 1000 values - change max for more data
  scale_x_continuous(
    breaks = seq(
      min(indicators_by_year_data$year), 
      max(indicators_by_year_data$year),
      by = 5)) +
  scale_fill_manual(
    values = dp_fill_palette,
    labels = to_title_case
  ) +
  labs(
    x = "Year",
    y = "Number of Indicators Represented",
    fill = "Dimension"
  ) +
  SMdocs::dp_theme

# Save to outputs
ggsave(
  indicator_plot,
  filename = '3_coverage/output/fig_indicator_coverage.png',
  width = 5,
  height = 4,
  units = 'in',
  dpi = 300,
  bg = 'white'
)

# Render to Quarto
indicator_plot
```

Same data but in a table for easy reporting:

```{r}
# Get totals by year:
totals <- indicators_by_year_data %>% 
  group_by(year) %>% 
  summarize(total_indicators = sum(indicator_count)) %>% 
  ungroup()
get_reactable(totals)
```

Breakdown by dimension:

```{r}
dim_totals <- indicators_by_year_data %>% 
  group_by(year, dimension) %>% 
  summarize(total_indicators = sum(indicator_count)) %>% 
  ungroup()
get_reactable(dim_totals)
```

### Metrics

Here we have metric representation by year - another graph made by Isabella. This one is focused on the metrics themselves, and provides a better look at update schedules and data availability over time.

```{r}
#| label: Metric representation each year
#| fig-height: 8
#| fig-width: 7
#| out-width: 80%
# Join based on "variable_name"
metrics_and_meta <- SMdocs::dp_metrics %>%
  left_join(SMdocs::dp_meta %>%
              select(variable_name,
                     dimension,
                     index,
                     indicator,
                     metric),
            by = "variable_name") 
# str(metrics_and_meta)

# Group dimension columns together in ascending order; filter out before 2000 
# and NAs
heatmap_easy_data <- metrics_and_meta %>%
  filter(!is.na(indicator)) %>%
  filter(year >= 2000, year < 2025) %>%
  distinct(indicator, dimension, year, metric)

heatmap_easy_data <- heatmap_easy_data %>%
  mutate(
    dimension = factor(dimension, levels = sort(unique(dimension)))
  ) %>%
  filter(!is.na(metric)) %>%
  arrange(dimension, metric) %>%
  mutate(
    metric = factor(metric, levels = rev(unique(metric)))
  )

heatmap_easy_data <- heatmap_easy_data %>%
  distinct(year, dimension, metric)

# Full grid of all combinations (to show blanks)
full_grid <- expand.grid(
  metric = levels(heatmap_easy_data$metric),
  year = unique(heatmap_easy_data$year),
  stringsAsFactors = FALSE
)

# Left join to get "present" boxes
data_full <- left_join(full_grid, heatmap_easy_data, by = c("metric", "year"))

# Make sure metric ordering is right
data_full <- data_full %>%
  mutate(
    metric = factor(
      metric,
      levels = levels(heatmap_easy_data$metric)
    )
  )

metric_heatmap <- ggplot(
  data_full, 
  aes(x = as.factor(year), y = metric, fill = dimension)
) +
  geom_tile(color = "gray80", linewidth = 0.25) +
  scale_y_discrete() +
  scale_x_discrete(
    breaks = seq(2000, 2024, 6) #
    # breaks = seq(2000, 2024, 4)
  ) +
  scale_fill_manual(
    values = c(dp_fill_palette, "FALSE" = "white"),
    na.value = "white",
    labels = to_title_case,
    na.translate = FALSE
  ) +
  labs(
    x = "Year",
    y = "Metric",
    fill = "Dimension"
  ) +
  SMdocs::dp_theme + 
  theme( 
    legend.position = 'right' #
  ) +
  # scale_color_discrete(guide = guide_legend(keywidth = 0.4, keyheight = 0.4)) +
  guides(
    # fill = guide_legend(nrow = 2, byrow = TRUE),
    fill = guide_legend(direction = 'vertical')
  ) +
  coord_fixed(ratio = 1.5) #

# Save to outputs for paper
ggsave(
  metric_heatmap,
  filename = '3_coverage/output/fig_metric_heatmap.png',
  width = 6,
  height = 8,
  units = 'in',
  dpi = 300,
  bg = 'white'
)

# Render to Quarto
metric_heatmap
```



```{r}
#| label: Indicator representation by year
#| include: false
# Join based on "variable_name"

## This is the same as the figure above, but for indicators instead of metrics.
# confusingly put in the metrics section
metrics_and_meta <- SMdocs::dp_metrics %>%
  left_join(SMdocs::dp_meta %>%
              select(variable_name,
                     dimension,
                     index,
                     indicator,
                     metric),
            by = "variable_name") 
str(metrics_and_meta)

# Group dimension columns together in ascending order; filter out before 2000 
# and NAs
heatmap_easy_data <- metrics_and_meta %>%
  filter(!is.na(indicator)) %>%
  filter(year >= 2000, year < 2025) %>%
  distinct(indicator, dimension, year, metric)

heatmap_easy_data <- heatmap_easy_data %>%
  mutate(
    dimension = factor(dimension, levels = sort(unique(dimension)))
  ) %>%
  filter(!is.na(metric)) %>%
  arrange(dimension, metric) %>%
  mutate(
    metric = factor(metric, levels = rev(unique(metric))),
    indicator = factor(indicator, levels = unique(indicator))
  )
str(heatmap_easy_data)

heatmap_easy_data <- heatmap_easy_data %>%
  distinct(year, dimension, indicator) %>%
  arrange(dimension, indicator)

# Full grid of all combinations (to show blanks)
full_grid <- expand.grid(
  indicator = levels(heatmap_easy_data$indicator),
  year = unique(heatmap_easy_data$year),
  stringsAsFactors = FALSE
)

# Left join to get "present" boxes
data_full <- left_join(full_grid, heatmap_easy_data, by = c("indicator", "year"))

# Ensure indicator maintains proper factor ordering from heatmap_easy_data
data_full <- data_full %>%
  mutate(
    indicator = factor(
      indicator, 
      levels = rev(levels(heatmap_easy_data$indicator))
    )
  )

# Save this for posterity
data_full %>% 
  arrange(dimension, year, indicator) %>% 
  write_csv('3_coverage/output/abstract_indicator_data.csv')

indicator_heatmap <- ggplot(
  data_full, 
  aes(x = as.factor(year), y = indicator, fill = dimension)
) +
  geom_tile(color = "gray80", linewidth = 0.25) +
  scale_y_discrete() +
  scale_x_discrete(
    breaks = seq(2000, 2024, 6) #
    # breaks = seq(2000, 2024, 4)
  ) +
  scale_fill_manual(
    values = c(dp_fill_palette, "FALSE" = "white"),
    na.value = "white",
    labels = to_title_case,
    na.translate = FALSE
  ) +
  labs(
    x = "Year",
    y = "Indicator",
    fill = "Dimension"
  ) +
  SMdocs::dp_theme + 
  theme( 
    legend.position = 'right' #
  ) +
  # scale_color_discrete(guide = guide_legend(keywidth = 0.4, keyheight = 0.4)) +
  guides(
    # fill = guide_legend(nrow = 2, byrow = TRUE),
    fill = guide_legend(direction = 'vertical')
  ) +
  coord_fixed(ratio = 1.5) #

ggsave(
  indicator_heatmap,
  filename = '3_coverage/output/fig_indicator_heatmap.png',
  width = 6,
  height = 8,
  units = 'in',
  dpi = 300,
  bg = 'white'
)

indicator_heatmap
```

