---
title: "Correlations"
format:
  html:
    fig-dpi: 200
    code-fold: true
editor_options: 
  chunk_output_type: console
warnings: false
execute:
  cache: false
---

```{r}
#| label: prep
#| cache: false
#| output: false
options(scipen = 999)

pacman::p_load(
  dplyr,
  htmltools,
  stringr,
  tidyr,
  reactable,
  purrr,
  broom,
  knitr,
  kableExtra,
  ggplot2,
  ggpubr,
  psych,
  tibble,
  Hmisc,
  reshape2,
  viridisLite,
  conflicted,
  readr,
  ggtext,
  snakecase
)

# GitHub packages
pacman::p_load_gh(
  'Food-Systems-Research-Institute/SMdata',
  'Food-Systems-Research-Institute/SMdocs'
)

conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  dplyr::pull(),
  dplyr::summarize(),
  stats::lag(),
  base::setdiff(),
  .quiet = TRUE
)
```

This page explores correlations among metrics across counties using the latest available time point for each metric.

## Wrangling

Here, filter our dataset to latest time points for each metric, filter to county data only, and join with a crosswalk that matches variable names to metric names. We then make an ordered framework that we use to arrange axes in a correlation matrix below.

```{r}
#| label: wrangle
#| output: false
#| fig-height: 10
#| fig-width: 10
# metric crosswalk
crosswalk <- SMdocs::dp_meta %>% 
  select(variable_name, metric)

# Get df ready
str(SMdata::metrics)
df <- SMdata::metrics %>% 
  filter(variable_name %in% SMdocs::dp_meta$variable_name) %>% 
  SMdocs::get_latest_year() %>% 
  SMdata::filter_fips('new') %>% 
  mutate(variable_name = str_split_i(variable_name, '_', 1)) %>% 
  left_join(crosswalk) %>% 
  select(-variable_name) %>% 
  pivot_wider(
    id_cols = fips,
    values_from = value,
    names_from = metric
  )
str(df)

# Get an ordered framework from which we will pull metrics in alphabetical order
# and get the counts of metrics per dimension that we will use to draw lines
# between dimensions
str(dp_meta)
ordered_framework <- dp_meta %>% 
  arrange(dimension, metric) %>% 
  filter(metric %in% names(df))
ordered_framework
```

## Figure

This figure shows absolute values of Spearman rank correlation coefficients among county level metrics. Green cells show significant correlations, while grey cells are insignificant. Darker green cells are stronger correlations and lighter green cells are weaker (but still significant).

Also, credits to Isabella and Tessa for making and improving this figure substantially.

```{r}
#| label: make_cor_table
#| output: false
# Get line placements to divide dimensions
# Reverse them to match alphabetical order of dimensions top to bottom
# Get cumulative sums to space them out across graph
# Then add 0.5 to each to put them in between cells
counts <- ordered_framework %>% 
  group_by(dimension) %>% 
  summarize(count = n()) %>% 
  pull(count)
hline_placements <- counts %>% 
  rev() %>% 
  cumsum() %>% 
  {. + 0.5}
hline_placements

vline_placements <- counts %>% 
  cumsum() %>% 
  {. + 0.5}
vline_placements

# Reorder our df in proper metric order
metric_order <- ordered_framework %>% 
  pull(metric)
df <- df %>% 
  select(fips, any_of(metric_order)) %>% 
  select(rev(everything()))
str(df)

# Create matrix of values
mat <- df %>% 
  na.omit() %>% 
  select(-fips) %>% 
  as.matrix()

# Get correlations
cor <- rcorr(mat, type = 'spearman')

# Melt correlation values and rename columns
cor_r <- reshape2::melt(cor$r) %>% 
  setNames(c('var_1', 'var_2', 'value'))

# Save p values
cor_p <- melt(cor$P)
p.value <- cor_p$value

# Absolute values of correlations
cor_r <- cor_r %>%
  mutate(value = abs(value))

# Pulling out significant p values
cor_r <- cor_r %>%
  mutate(p_value = ifelse(var_1 == var_2, 1, cor_p$value)) %>% # Make diagonal line non-significant
  mutate(significant = ifelse(p_value <= 0.05, TRUE, FALSE))

# Reverse the order to use for x-axis labels (start at Economics)
reversed_metric_order <- rev(metric_order)

# Create a reversed version of the cor_r data frame for the x-axis
cor_r <- cor_r %>%
  mutate(
    var_1 = factor(var_1, levels = reversed_metric_order)
  )

# Remove last cols []
str(cor_r)
cor_r <- cor_r %>%
  filter(var_1 != var_2)
  # filter(!(var_1 == "Average Weekly Wages" & var_2 == "Violent Crime Rate")) %>%
  # filter(!(var_1 == "Violent Crime Rate" & var_2 == "Average Weekly Wages"))
str(cor_r)
  
# Only set values for lower triangle of the matrix
cor_r <- cor_r %>%
  mutate(
    row_idx = as.integer(factor(var_1, levels = reversed_metric_order)),
    col_idx = as.integer(factor(var_2, levels = reversed_metric_order)),
    value = ifelse(row_idx <= col_idx, NA, value)  # Set upper triangle to NA
  ) %>%
  select(-row_idx, -col_idx)  # Remove temporary columns

# Pasting dimension colors to metrics
colored_labels <- ordered_framework %>%
  mutate(
    label = paste0(
      "<span style='color:",
      dp_text_palette[dimension],
      "'>",
      to_title_case(str_replace_all(metric, "_", " ")),
      "</span>"
    )
  ) %>%
  select(metric, label) %>%
  deframe()

## Removing topmost y axis label and rightmost x axis label
# Create modified label vectors
x_labels <- colored_labels
x_labels[metric_order[length(metric_order)]] <- ""  # remove last x-axis label
y_labels <- colored_labels
y_labels[reversed_metric_order[length(reversed_metric_order)]] <- "" # Blank topmost y-axis label

# Create tick vectors (TRUE = show tick, FALSE = hide tick)
x_ticks <- rep(TRUE, length(metric_order))
x_ticks[length(metric_order)] <- FALSE  # Hide rightmost x-axis tick

y_ticks <- rep(TRUE, length(reversed_metric_order))
y_ticks[length(reversed_metric_order)] <- FALSE  # Hide topmost y-axis tick

# Dummy data frame for dimension legend
dimension_legend_df <- tibble(
  x = -Inf,
  y = -Inf,
  dimension = factor(names(dp_text_palette), levels = names(dp_text_palette))
)

# Make heatmap
corplot <- ggplot(cor_r, aes(x = var_1, y = var_2)) +
  
  # Color in significant correlations
  geom_tile(
    data = cor_r %>% filter(!is.na(value) & significant), 
    aes(fill = value)
  ) +
  
  # Make non-significant correlations white
  geom_tile(
    data = cor_r %>% filter(!is.na(value) & !significant), 
    fill = "gray90"
  ) +
  
  # Grid lines for lower triangle
  geom_tile(
    data = cor_r %>% filter(!is.na(value)),
    color = "gray60",
    linewidth = 0.2,
    fill = NA
  ) +

  geom_point(
    data = dimension_legend_df,
    aes(x = x, y = y, color = dimension),
    shape = 15,
    size = 5,
    inherit.aes = FALSE,
    show.legend = TRUE,
    alpha = 0  # fully transparent points, so they don't show up
  ) +
  
  scale_fill_gradient2(
    low = "white", 
    high = "#1b7837", 
    midpoint = 0,
    name = "Correlation",
    na.value = "transparent"
  ) +
  scale_color_manual(
    values = dp_text_palette,
    labels = str_to_title(names(dp_text_palette)),
    name = "Dimension"
  ) +
  
  # Legend overrides
  guides(
    fill = guide_colorbar(
    title = "Correlation", 
    order = 1
  ),
    color = guide_legend(
      override.aes = list(shape = 15, size = 5, alpha = 1),
      nrow = 2,
      order = 2
    )
  ) +
  
  # Horizontal Lines
  geom_segment(
    x = 0,
    xend = vline_placements[4],
    y = hline_placements[1], 
    yend = hline_placements[1]
  ) +
  geom_segment(
    x = 0,
    xend = vline_placements[3],
    y = hline_placements[2], 
    yend = hline_placements[2]
  ) +
  geom_segment(
    x = 0,
    xend = vline_placements[2],
    y = hline_placements[3], 
    yend = hline_placements[3]
  ) +
  geom_segment(
    x = 0,
    xend = vline_placements[1],
    y = hline_placements[4], 
    yend = hline_placements[4]
  ) +
    
  # Vertical Lines
  geom_segment(
    x = vline_placements[1],
    xend = vline_placements[1],
    y = 0, 
    yend = hline_placements[4] 
  ) +
  geom_segment(
    x = vline_placements[2],
    xend = vline_placements[2],
    y = 0, 
    yend = hline_placements[3] 
  ) +
  geom_segment(
    x = vline_placements[3],
    xend = vline_placements[3],
    y = 0, 
    yend = hline_placements[2] 
  ) +
  geom_segment(
    x = vline_placements[4],
    xend = vline_placements[4],
    y = 0, 
    yend = hline_placements[1] 
  ) +
    
  # Correct x and y axis labels with the order
  scale_x_discrete(
    labels = x_labels,
    limits = metric_order
  ) +
  scale_y_discrete(
    labels = y_labels,
    limits = reversed_metric_order
  ) +
  
  theme(
    axis.text.x = element_markdown(angle = 45, hjust = 1),
    axis.text.y = element_markdown(),
    axis.text = element_text(family = "Times New Roman"),
    axis.title = element_blank(),
    axis.ticks.length.x = unit(ifelse(x_ticks, 0.15, 0), "cm"),
    axis.ticks.length.y = unit(ifelse(y_ticks, 0.15, 0), "cm"),
    plot.title = element_text(hjust = 0),
    legend.position = "top",
    legend.title.position = "top",
    legend.justification = "center",
    legend.title = element_text(
      size = 10,
      family = "Times New Roman",
      hjust = 0.5
    ),
    legend.text = element_text(size = 9, family = "Times New Roman"),
    legend.key = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
    panel.background = element_rect(fill = "transparent", colour = NA)
  )

ggsave(
  plot = corplot,
  filename = '4_correlations/output/fig_corplot.png',
  height = 9,
  width = 9,
  dpi = 300,
  bg = 'white'
)
```


```{r}
#| label: show_cor_table
#| echo: false
#| fig-height: 9
#| fig-width: 9
corplot
```

Reading the tea leaves here, it looks like we have lots of strong correlations within the Health dimension. This might suggest that the health metrics are particularly cohesive. It might also suggest that they are redundant, and we could use fewer metrics to measure the dimension. In contrast, the Economics dimension, for example, has very few correlations within its metrics. Again, this might signify a lack of theoretical coherence among metrics or that it is indeed a diverse dimension that requires disparate data sources to measure.

As for significant inter-dimension correlations, there are surprisingly few. This is probably a good thing on average, as strong correlations between dimensions would mean double counting of similar constructs. On the other hand, we might want to see more of these if we are to expect that changes in one dimension cause subsequent changes in another. For example, stricter environmental policies might lead to better environmental outcomes, but we would also want to see any changes it causes in the Production and Economic dimensions.

## LaTeX Table

Making $\LaTeX$ table for paper.

```{r}
#| label: latex_table
#| output: false
# Latex table
head(cor_r)
head(cor_p)
df <- full_join(cor_r, cor_p, by = join_by(var_1 == Var1, var_2 == Var2)) %>% 
  select(1:4) %>% 
  setNames(c('var_1', 'var_2', 'r', 'p'))
str(df)

# Format and add stars
df <- df %>% 
  filter(!is.na(r)) %>% 
  mutate(
    stars = case_when(
      p < 0.001 ~ '***',
      p > 0.001 & p < 0.01~ '***',
      p > 0.01 & p < 0.05 ~ '**',
      .default = ''
    ),
    r = format(round(r, 3), nsmall = 3),
    r = paste0(r, stars)
  ) %>% 
  select(-c(stars, p))
str(df)

# Pivot
df <- df %>%
  pivot_wider(
    names_from = var_2,
    values_from = r,
    id_cols = var_1
  )
str(df)

# Make names short, remove var header, turn NA into ''
truncate_strings <- function(x) {
  if (str_length(x) > 20) {
    paste0(str_sub(x, end = 20), '...')
  } else {
    x
  }
}
df <- df %>% 
  mutate(
    var_1 = as.character(var_1),
    var_1 = map_vec(var_1, truncate_strings),
    across(!var_1, ~ ifelse(is.na(.x), '', .x))
  ) %>%
  setNames(c(map(names(.), truncate_strings)))
str(df)
colnames(df)[1] <- ' '
str(df)

# Paste rotatebox around column names
colnames(df)[2:ncol(df)] <- paste0(
  '\\makecell{\\rotatebox{90}{',
  colnames(df)[2:ncol(df)],
  '}}'
)

df %>% 
  kbl(
    format = "latex",
    caption = 'Correlation matrix of county level metrics',
    label = 'tab_correlations',
    booktabs = TRUE,
    escape = FALSE
    # align = 'c'
  ) %>%
  kable_styling(
    font_size = 8,
    bootstrap_options = c(
      "condensed"
    ),
    latex_options = 'scale_down'
  ) %>% 
  landscape() %>% 
  footnote(
    general_title = "",
    general = 
      '\\\\textit{Note: }
    Spearman correlations among county level metrics using latest time points
    available. Connecticut counties are not included in analyses.
    ***: p < 0.001, **: p < 0.01, *: p < 0.05. ',
    threeparttable = TRUE,
    escape = FALSE
  ) %>% 
  save_kable(
    file = '4_correlations/output/tab_correlations.tex'
  )
```

## Long Table

Also including a long-format interactive table that shows p-values and coefficients. Use this for reporting in paper.

```{r}
#| label: cor_table_long
# Create long table with p values and stars
tab <- cor_r %>% 
  rename(rho = value, p = p_value) %>% 
  filter(!is.na(rho)) %>% 
  mutate(sig = ifelse(p < 0.05, '*', ''))

tab %>% 
  get_reactable(
    defaultColDef = colDef(
      format = colFormat(digits = 3)
    )
  )
```

Reporting for paper: 

```{r}
#| output: false
# Join dimensions back to correlations for reporting
tree <- SMdocs::dp_tree %>% 
  select(dimension, metric) %>% 
  filter(str_detect(metric, 'NONE', negate = TRUE))

# How many correlations are significant overall
(perc_sig <- round(mean(tab$sig == '*') * 100, 2))
(perc_unsig <- 100 - perc_sig)

# Internal correlations among each dimension
dimensions <- unique(tree$dimension)
internal_cors <- map(dimensions, ~ {
  dim_metrics <- tree %>% 
    filter(dimension == .x) %>% 
    pull(metric)
  dim_cors <- tab %>% 
    filter(var_1 %in% dim_metrics & var_2 %in% dim_metrics) %>% 
    pull(sig)
  dim_sig_cors <- round(mean(dim_cors == '*') * 100, 2)
}) %>% 
  setNames(c(dimensions)) %>% 
  as.data.frame()
internal_cors

# Check percentage of sig correlations for a couple of examples in paper:
str(tab)
vars <- c(
  'Violent Crime Rate',
  'Land and Building Value per Acre'
)
examples <- map(vars, ~ {
  sigs <- tab %>% 
    filter(var_1 == .x | var_2 == .x) %>% 
    pull(sig)
  round(mean(sigs == '*') * 100, 2)
}) %>% 
  setNames(c(vars)) %>% 
  bind_rows()
examples
```

Percentage of significant internal correlations in each dimension:

```{r}
internal_cors
```

